---
title: "DNTransect_Summarizing"
author: "Amy Henry"
date: "2023-01-24"
output: html_document
---

```{r}
library(tidyverse)
library(vegan)

load(file = "DNTransect_Primary.Rdata")
```


## Species Richness
Calculate species richness for each TRANSECT for each Treatment and Seed Source combination

-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}

```

## Shannon Diversity
Calculate Shannon Diversity for each TRANSECT for each Treatment and Seed Source combination

-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}

```


## Calculate Percent Cover with PRESENCE/ABSENCE
Convert Hits column to 0/1 (i.e., presence/absence), then sum for percent cover for each TRANSECT for each Treatment and Seed Source combination

-- For each Focal Species (ENCCAL, ERIFAS, ARTCAL, SALMEL)
-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}

```

## Calculate Rough Abundance with COUNTS OF HITS
Sum the count of hits for each TRANSECT for each Treatment and Seed Source combination

-- For each Focal Species (ENCCAL, ERIFAS, ARTCAL, SALMEL)
-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}

```



