---
title: "DNTransect_Summarizing"
author: "Amy Henry"
date: "2023-01-24"
output: html_document
---

```{r}
library(tidyverse)
library(vegan)
library(lubridate)

load(file = "DNTransect_Primary.Rdata")
```

Steps:
1. Create an Index column for all transects
2. Create one column per species using pivot_wider()
3. Separate the Metadata colu,ns from the species data columns, keeping Index in common
4. make the species data columnds a matrix to use functions from the `vegan` package
5. Merge index information baack to your calculated matrix

## Species Richness
Calculate species richness for each TRANSECT for each Treatment and Seed Source combination

-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}
#Creating your index
DN_index <- DNTransect_Primary %>% 
  mutate(Index = paste(year(Date),Block,Plot,Transect, sep = "-"))

DN_Hits_Matrix <- DN_index %>% 
  filter(Type =="Plant") %>%
  select(Index, SpeciesCode, Hits) %>% 
  group_by(Index, SpeciesCode) %>%
  summarize(Hits = sum(Hits)) %>%
  pivot_wider(names_from = SpeciesCode, values_from = Hits) %>%
  replace(is.na(.), 0)

DN_PresABS_Matrix <- DN_index %>% 
  filter(Type == "Plant") %>%
  select(Index, SpeciesCode, Hits) %>%
  group_by(Index, SpeciesCode) %>%
  summarize(Hits = sum(Hits)) %>%
  mutate(PresAbs = case_when(
    Hits >= 1 ~ 1,
    Hits < 1 ~ 0
  )) %>%
  pivot_wider(names_from = SpeciesCode, values_from = Hits) %>%
  replace(is.na(.), 0)

DN_diversity <- DN_PresABS_Matrix %>%
  mutate(SpRich = rowSums(across(where(is.numeric)), na.rm=TRUE))

view(DN_diversity)

Transect_Index <- DN_index %>% 
  select(Date,Block:Transect, Index) %>% 
  distinct()

DN_diversity_index <- DN_diversity %>% 
  select(Index, SpRich) %>% 
  full_join(Transect_Index) %>%
  select(Index, Date:Transect, SpRich)

DN_diversity_index %>% ggplot() +
  aes(x = Treatment, y = SpRich, fill = Region) +
  geom_boxplot()

```

## Shannon Diversity
Calculate Shannon Diversity for each TRANSECT for each Treatment and Seed Source combination

-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}
Shannon <- DN_Hits_Matrix %>% 
  ungroup() %>%
  select(AVEFAT:SILSPP) %>% 
  diversity(index = "shannon", MARGIN = 1 )

DN_Hits_Matrix %>% 
  select(Index) %>% 
  bind_cols(Shannon) %>%
  rename(H = `...2`)

DN_diversity_index <- DN_diversity_index %>% full_join(Shannon2)

DN_diversity_index %>% ggplot() + 
  aes(x = Treatment, y = H, fill = Region) +
  geom_boxplot()

```


## Calculate Percent Cover with PRESENCE/ABSENCE
Convert Hits column to 0/1 (i.e., presence/absence), then sum for percent cover for each TRANSECT for each Treatment and Seed Source combination

-- For each Focal Species (ENCCAL, ERIFAS, ARTCAL, SALMEL)
-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}

```

## Calculate Rough Abundance with COUNTS OF HITS
Sum the count of hits for each TRANSECT for each Treatment and Seed Source combination

-- For each Focal Species (ENCCAL, ERIFAS, ARTCAL, SALMEL)
-- Total
-- Native
-- Non-native
-- LifeForm
---Native
---Non-native
```{r}

```



