---
title: "test_github.Rmd"
output: html_document
date: "2023-01-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#Install Packages

library(tidyverse)
library(readxl)
library(janitor)
library(vegan)
library(lubridate)
library(dataspice)
```


Phase I: Data Cleaning 

```{r}
Spring_2020_raw_test <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)
Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE) 
Spring_2021_raw <- read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)
Spring_2022_raw <- read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

```


```{r}
Spring_data_raw_pre <- bind_rows(Spring_2020_raw, Spring_2021_raw, Spring_2022_raw)
Spring_data_raw_pre <- Spring_data_raw_pre %>% 
 ## row_to_names(row_number(1)) %>% ## change the first row to be the name of the column
  rename(Transect = `Transect #`, Hits = `# Hits`)%>%
  filter(Date != "Date" & Date !="mm/dd/yy")

nrow(Spring_data_raw_pre)
```


```{r}
Spring_data_raw <- Spring_data_raw_pre %>%
  mutate(Date = as.Date(as.numeric(Date), origin = "1899-12-30"))%>%
  mutate(Distance = as.numeric(Distance), 
         Block = as.factor(Block), 
         Plot = as.factor(Plot), 
         Region = as.factor(Region), 
         Treatment = as.factor(Treatment),
         Transect = as.factor(Transect),
         Hits = as.numeric(Hits))%>%
  filter(is.na(Date) == FALSE)

str(Spring_data_raw)
```


```{r}
Full_Species_List_Glossary_reference_only_ <- read_excel("Full Species List Glossary (reference only).xlsx")

SpeciesKey <- Full_Species_List_Glossary_reference_only_%>%
  select(1:7)%>%
  rename(SpeciesCode =`Species Code (A-Z)`, 
         NativeStatus = `Native/Non-Native`, 
         CommonName = `Common Name`, 
         SciName = `Full Scientific Name (Genus species)`, 
         AltSciName = `other/old scientific names`)

unique(Spring_data_raw$Code)%>% write.csv(file = "DNCodes_raw.csv")
DNCodes <- read_csv("DNCodes.csv")


```


```{r}
intermediate <- Spring_data_raw%>%
  full_join(DNCodes, by = "Code")%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  mutate(LifeForm = case_when(is.na(LifeForm) == FALSE ~ LifeForm, TRUE ~ as.character(Lifeform)),
         CommonName = case_when(
           SpeciesCode == "L" ~ "littter",
           SpeciesCode == "B" ~ "bare ground",
           SpeciesCode == "W" ~ "coarse woody debris",
           SpeciesCode == "F" ~ "fine debris",
           SpeciesCode == "R" ~ "rock",
           SpeciesCode == "S" ~ "stem",
           SpeciesCode == "T" ~ "thatch",
           TRUE ~ as.character(CommonName)
         ),
         SciName = case_when(
           Type == "Ground" ~ CommonName,
           TRUE ~ as.character(SciName)
         )
           
)%>%
  select(-Code, -Lifeform )%>%
  fill(Transect)%>% # will copy the value from the row above 
  mutate(Hits = case_when(Type == "Plant" ~ 1,
                          Type == "Ground"~ 1))%>% #Change NA hits to = 1 hit when type = plant, ground = 0/NA
  rename(Plant_Family = `Plant Family`)

HITS <- intermediate%>%
  filter(is.na(Hits) == TRUE)

#Add column for ground cover?

#view(HITS)

#testing for unique values for each column
unique(intermediate$Treatment)
unique(intermediate$Plot)
unique(intermediate$Block)
unique(intermediate$Region)
unique(intermediate$Transect)
unique(intermediate$Volunteer) ## maybe need to change volunteer to "Y" and NA to "N" ?
unique(intermediate$Distance) ## not sure if having 52.0 is normal
unique(intermediate$AltSciName)
unique(intermediate$SciName)
unique(intermediate$CommonName) ## number of total SciName and CommonName are not matching: Avena fatua and Avena barbata are both wild oat. Maybe need to differentiate in the common name?
SciCommon <- intermediate%>%
  filter(SciName == "Avena barbata")

#figure the number of rows of ground and NAs Type, will be used to see if all the NAs in status, native status, and Plant_Family matches (Asumming all Plant Type data will not have these columns missing)
test_na <- intermediate %>%
  filter(Type == "Ground" | is.na(Type) == TRUE)
nrow(test_na)


unique(intermediate$Status)
Status <- intermediate%>%
  filter(is.na(Status) == TRUE)
nrow(Status)

unique(intermediate$NativeStatus)
Native <- intermediate%>%
  filter(is.na(NativeStatus) == TRUE)
nrow(Native)

unique(intermediate$Plant_Family)
PlantFamily <- intermediate%>%
  filter(is.na(Plant_Family) == TRUE)
nrow(PlantFamily)

#numbers of rows of plant family and Status do not match, filter the value out to see what plant species is missing information
unique(intermediate$NativeStatus)
Native <- intermediate%>%
  filter(is.na(NativeStatus) == TRUE & Type != "Ground" & is.na(Type) == FALSE) 
unique(Native$SpeciesCode)
unique(intermediate$Plant_Family)
PlantFamily <- intermediate%>%
  filter(is.na(Plant_Family) == TRUE & Type != "Ground" & is.na(Type) == FALSE)
unique(PlantFamily$SpeciesCode)
#AVESPP = non-native, BROSPP, SILSPP can be native or non-native, FESMYE = FESMYU, EROSPP = non-native, CONBON, PHAMIN =  not found in the full species list, need to check the species
```


## Create Metadata file (csv)
```{r}

# Create separate dataframe to convert to csv 

Metadata <- intermediate %>% 
  colnames(.) 

# View columns to write descriptions for
Metadata

#Write out descriptions of columns in their exact order) Note: Still need units but am unsure. Will consult Moises next class
Description <- c(
  'Collection day (YYYY-MM-DD)',
  'Data collector names',
  'Block number (1-4)',
  'Plot number (1-9)',
  'Water treatment type (I = Irrigation, A = Ambient, E = Exclusion)',
  'Seed source (M = Mesic, X = Xeric)',
  'Transect number (1-5)', 
  'Collection point distance from start of transect', ## from 0.5-5 and every .5 meter?
  'How many times a plant touches the collection pole',
  'Other notes',
  'Species ID code (6-letters)',
  'Classification of groundcover (Plant or Ground)',
  'Plant Status (Dead or alive)', #unsure about this description
  'Type of plant (Shrub, Herb, Grass, Forb)',
  'Is it planted or self seeded',
  'Scientific name',
  'Is it native or non native',
  'Species common name',
  'Other or Alternative Scientific names',
  'Plant family'
  )

# Combine column name dataset with newly made column
Metadata_better <- cbind(Metadata, Description)


#check new dataframe
view(Metadata_better)






# Convert to csv. Make sure to change the pathway to be for your computer
write.csv(Metadata_better, "Dataset_Metadata_20230201.csv", row.names = TRUE)


# Check csv file
read.csv(file = "Dataset_Metadata_20230201.csv")  




# Example codes from the internet
#tmp <- read.csv("original_file.csv")
#tmp <- cbind(tmp, new_column)
#write.csv(tmp, "modified_file.csv")
```

## export data to .csv and Rdata
```{r}
write_csv(intermediate, "Dataset_PRIMARY_20230124.csv")
saveRDS(intermediate, file = "Dataset_PRIMARY_20230124.Rdata")

#Amy's code to save and turn in
#save(DNTransect_Primary, file = " ")
```


Phase II Analysis 

##Importing data directly from google drive
```{r}
library (googlesheets4)

dat2020 <- read_sheet

```

## Next steps for class. Amy will release full code file soon.
# One student for each calculation

#Parameters to calculate x Treatment x Seed Source
Species richness
total 
Native
Non-native
Lifeform
Native


#Shannon Diversity
Total
Native
non -native
Lifeform
Native
Non-native
Percent cover


#Focal species (ENCCAL, ERIFAS, ARTCAL, SALMEL)
Total
Native
non -native
Lifeform
Native
Non-native


##percent cover
point/total points

## Cover dominace
hits/total points (hits per point)

# occupancy
points/9
# abundacne 
hits/9
# relative abundance 
hits native/ hits total


## Formatting the data for calculation and GLM (percent vover)
```{r}
load(file = "DNTransect_Primary.Rdata")

DN_index <- DNTransect_Primary %>% 
  mutate(Index = paste(year(Date),Block,Plot,Transect, sep = "-"))

SpeciesPerCover <- DN_index %>%
  filter(Type == "Plant")%>%
  select(Index, SpeciesCode, Hits)%>%
  mutate(PresAbs = case_when(
    Hits >= 1 ~ 1,
    Hits < 1 ~ 0
    )) %>%
  group_by(Index, SpeciesCode)%>%
  summarise(PresAbsSum = sum(PresAbs))%>%
  mutate(PerCover = PresAbsSum/9) %>%
  pivot_wider(names_from = SpeciesCode, values_from = PerCover)%>%
  replace(is.na(.), 0)
  
```

```{r}
NativePerCover <- SpeciesPerCover %>%
  pivot_longer(3:ncol(SpeciesPerCover), names_to = "SpeciesCode", values_to = "PerCover")%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  select(-PresAbsSum)%>%
  filter(NativeStatus == "Native")%>%
  group_by(Index)%>%
  summarise(NativePerCover = sum(PerCover))

```


```{r}
DN_diversity_index_Native <- DN_diversity_index%>% left_join(NativePerCover)

ggplot(DN_diversity_index_Native)+
  aes(y = NativePerCover, x = Treatment, fill = Region)+
  geom_boxplot()
```
## creating histogram to visulaise data distribution
```{r}
DN_diversity_index_Native%>%
  filter(Region != "C")%>%
  ggplot()+
  aes(x = NativePerCover)+
  geom_histogram()+
  facet_grid(Treatment~Region)


# Three components of GLM
# 1. random component
# 2. systematic component
# 3. link function
library(lme4)
library(lmerTest)

test_1 <- DN_diversity_index_Native %>% 
  mutate (Year = year(Date))%>%
  filter(Region != "C", Year == 2022)

model <- lmer(NativePerCover~ Treatment * Region + (1| Block),
             data = test_1
             #family = "gaussian"
             )

model_noRE <- lm(NativePerCover~ Treatment * Region,  
                 data = test_1)
summary(model)

plot(model)

# To read the rsult -
##Estimate - 
###Treatment E and I (*) compared to ambient
###Region X compared to Region M (*)

#Akaike Information Criterion
## smaller values = better fit model

AIC(model) # model with random effect of block
AIC(model_noRE) # model with no block

```


STEPS:
1. Calculate your variable
2. Add it to the table of diversity indices (which includes transect info)
3. Visualize it using box plot and/or histogram
4. Choose your random component
5. Compose your systematic component (your formula)
6. Decide if you have random effects to include
7. Choose which modeling function to use in R
- lm for normal fixed effects only
- glm for non-normal, fixed effects only
- lmer for normal with random effects
- glmer for non-normal with random effects
8. Plot(model) to see residual plot
9. Examine fixed effects estimates in summary(model)
10. Compare fit of model to other possible modeling using AIC()


#For assignment 
One for biodiversity, one for percent cover

1. Total Native hits/transect - poisson
2. Total Native hits/ Total Hits (per transect) --> proportion -  Beta/binomial
3. PerCoverENCCAL (Pre-abs/9) - binomial
4. Total Non-native hits/transect - poisson
5. 
6. 

# Calculating the variable and the merge it with diversity index
```{r}
NonNative_hits <- DN_index %>%
  filter(Type == "Plant")%>%
  select(Index, SpeciesCode, Hits)%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  filter(NativeStatus == "Non-Native")%>%
  group_by(Index)%>%
  summarise(Non_Native_Hits = sum(Hits))%>%
  left_join(DN_diversity_index)

```

# Create histogram
```{r}
ggplot(NonNative_hits)+
  aes(y = Non_Native_Hits, x = Treatment, fill = Region)+
  geom_boxplot()

NonNative <- NonNative_hits%>%
  filter(Region != "C")

ggplot(NonNative)+
  aes(x = Non_Native_Hits)+
  geom_histogram()+
  facet_grid(Treatment~Region)
```


```{r}
NonNative_1 <- NonNative %>% 
  mutate (Year = year(Date))%>%
  filter(Region != "C", Year == 2022)

model_nonnative <- lmer(Non_Native_Hits~ Treatment * Region + (1| Block),
             data = NonNative_1
             #family = "gaussian"
             )

model_noRE_nonnative <- lm(Non_Native_Hits~ Treatment * Region,  
                 data = NonNative_1)
summary(model_nonnative)

plot(model_nonnative)

AIC(model_nonnative) 
AIC(model_noRE_nonnative) 
```

