---
title: "Noreen's Workspace"
output: html_document
date: "2023-02-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#Install Packages

library(tidyverse)
library(readxl)
library(janitor)
library(vegan)
library(lubridate)
library(dataspice)
```


## Phase I: Data Cleaning 
```{r}
Spring_2020_raw_test <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)
Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE) 
Spring_2021_raw <- read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)
Spring_2022_raw <- read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

```


```{r}
Spring_data_raw_pre <- bind_rows(Spring_2020_raw, Spring_2021_raw, Spring_2022_raw)
Spring_data_raw_pre <- Spring_data_raw_pre %>% 
 ## row_to_names(row_number(1)) %>% ## change the first row to be the name of the column
  rename(Transect = `Transect #`, Hits = `# Hits`)%>%
  filter(Date != "Date" & Date !="mm/dd/yy")

nrow(Spring_data_raw_pre)
```


```{r}
Spring_data_raw <- Spring_data_raw_pre %>%
  mutate(Date = as.Date(as.numeric(Date), origin = "1899-12-30"))%>%
  mutate(Distance = as.numeric(Distance), 
         Block = as.factor(Block), 
         Plot = as.factor(Plot), 
         Region = as.factor(Region), 
         Treatment = as.factor(Treatment),
         Transect = as.factor(Transect),
         Hits = as.numeric(Hits))%>%
  filter(is.na(Date) == FALSE)

str(Spring_data_raw)
```


```{r}
Full_Species_List_Glossary_reference_only_ <- read_excel("Full Species List Glossary (reference only).xlsx")

SpeciesKey <- Full_Species_List_Glossary_reference_only_%>%
  select(1:7)%>%
  rename(SpeciesCode =`Species Code (A-Z)`, 
         NativeStatus = `Native/Non-Native`, 
         CommonName = `Common Name`, 
         SciName = `Full Scientific Name (Genus species)`, 
         AltSciName = `other/old scientific names`)

unique(Spring_data_raw$Code)%>% write.csv(file = "DNCodes_raw.csv")
DNCodes <- read_csv("DNCodes.csv")


```


```{r}
intermediate <- Spring_data_raw%>%
  full_join(DNCodes, by = "Code")%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  mutate(LifeForm = case_when(is.na(LifeForm) == FALSE ~ LifeForm, TRUE ~ as.character(Lifeform)),
         CommonName = case_when(
           SpeciesCode == "L" ~ "littter",
           SpeciesCode == "B" ~ "bare ground",
           SpeciesCode == "W" ~ "coarse woody debris",
           SpeciesCode == "F" ~ "fine debris",
           SpeciesCode == "R" ~ "rock",
           SpeciesCode == "S" ~ "stem",
           SpeciesCode == "T" ~ "thatch",
           TRUE ~ as.character(CommonName)
         ),
         SciName = case_when(
           Type == "Ground" ~ CommonName,
           TRUE ~ as.character(SciName)
         )
           
)%>%
  select(-Code, -Lifeform )%>%
  fill(Transect)%>% # will copy the value from the row above 
  mutate(Hits = case_when(Type == "Plant" ~ 1,
                          Type == "Ground"~ 1))%>% #Change NA hits to = 1 hit when type = plant, ground = 0/NA
  rename(Plant_Family = `Plant Family`)

HITS <- intermediate%>%
  filter(is.na(Hits) == TRUE)

#Add column for ground cover?

#view(HITS)

#testing for unique values for each column
unique(intermediate$Treatment)
unique(intermediate$Plot)
unique(intermediate$Block)
unique(intermediate$Region)
unique(intermediate$Transect)
unique(intermediate$Volunteer) ## maybe need to change volunteer to "Y" and NA to "N" ?
unique(intermediate$Distance) ## not sure if having 52.0 is normal
unique(intermediate$AltSciName)
unique(intermediate$SciName)
unique(intermediate$CommonName) ## number of total SciName and CommonName are not matching: Avena fatua and Avena barbata are both wild oat. Maybe need to differentiate in the common name?
SciCommon <- intermediate%>%
  filter(SciName == "Avena barbata")

#figure the number of rows of ground and NAs Type, will be used to see if all the NAs in status, native status, and Plant_Family matches (Asumming all Plant Type data will not have these columns missing)
test_na <- intermediate %>%
  filter(Type == "Ground" | is.na(Type) == TRUE)
nrow(test_na)


unique(intermediate$Status)
Status <- intermediate%>%
  filter(is.na(Status) == TRUE)
nrow(Status)

unique(intermediate$NativeStatus)
Native <- intermediate%>%
  filter(is.na(NativeStatus) == TRUE)
nrow(Native)

unique(intermediate$Plant_Family)
PlantFamily <- intermediate%>%
  filter(is.na(Plant_Family) == TRUE)
nrow(PlantFamily)

#numbers of rows of plant family and Status do not match, filter the value out to see what plant species is missing information
unique(intermediate$NativeStatus)
Native <- intermediate%>%
  filter(is.na(NativeStatus) == TRUE & Type != "Ground" & is.na(Type) == FALSE) 
unique(Native$SpeciesCode)
unique(intermediate$Plant_Family)
PlantFamily <- intermediate%>%
  filter(is.na(Plant_Family) == TRUE & Type != "Ground" & is.na(Type) == FALSE)
unique(PlantFamily$SpeciesCode)
#AVESPP = non-native, BROSPP, SILSPP can be native or non-native, FESMYE = FESMYU, EROSPP = non-native, CONBON, PHAMIN =  not found in the full species list, need to check the species
```





##[Phase II Analysis] - I'm assigned PerCover ENCAL (Pres/Abs (0/1)/9(points)) Binomial





## Next steps for class. Amy will release full code file soon.
# One student for each calculation

#Parameters to calculate x Treatment x Seed Source
Species richness
total 
Native
Non-native
Lifeform
Native


#Shannon Diversity
Total
Native
non -native
Lifeform
Native
Non-native
Percent cover


#Focal species (ENCCAL, ERIFAS, ARTCAL, SALMEL)
Total
Native
non -native
Lifeform
Native
Non-native


##percent cover
point/total points

## Cover dominace
hits/total points (hits per point)

# occupancy
points/9
# abundacne 
hits/9
# relative abundance 
hits native/ hits total


## Formatting the data for calculation and GLM (percent vover)
```{r}
load(file = "DNTransect_Primary.Rdata")

DN_index <- DNTransect_Primary %>% 
  mutate(Index = paste(year(Date),Block,Plot,Transect, sep = "-"))

SpeciesPerCover <- DN_index %>%
  filter(Type == "Plant")%>%
  select(Index, SpeciesCode, Hits)%>%
  mutate(PresAbs = case_when(
    Hits >= 1 ~ 1,
    Hits < 1 ~ 0
    )) %>%
  group_by(Index, SpeciesCode)%>%
  summarise(PresAbsSum = sum(PresAbs))%>%
  mutate(PerCover = PresAbsSum/9) %>%
  pivot_wider(names_from = SpeciesCode, values_from = PerCover)%>%
  replace(is.na(.), 0)
  
```

```{r}
NativePerCover <- SpeciesPerCover %>%
  pivot_longer(3:ncol(SpeciesPerCover), names_to = "SpeciesCode", values_to = "PerCover")%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  select(-PresAbsSum)%>%
  filter(NativeStatus == "Native")%>%
  group_by(Index)%>%
  summarise(NativePerCover = sum(PerCover))

```


```{r}
DN_diversity_index <- DN_diversity_index%>% left_join(NativePerCover)

ggplot(DN_diversity_index)+
  aes(y = NativePerCover, x = Treatment, fill = Region)+
  geom_boxplot()
```
## creating histogram to visulaise data distribution
```{r}
DN_diversity_index%>%
  filter(Region != "C")%>%
  ggplot()+
  aes(x = NativePerCover)+
  geom_histogram()+
  facet_grid(Treatment~Region)


# Three components of GLM
# 1. random component
# 2. systematic component
# 3. link function
library(lme4)
library(lmerTest)

test_1 <- DN_diversity_index %>% 
  mutate (Year = year(Date))%>%
  filter(Region != "C", Year == 2022)

model <- lmer(NativePerCover~ Treatment * Region + (1| Block),
             data = test_1
             #family = "gaussian"
             )

model_noRE <- lm(NativePerCover~ Treatment * Region,  
                 data = test_1)
summary(model)

plot(model)

# To read the rsult -
##Estimate - 
###Treatment E and I (*) compared to ambient
###Region X compared to Region M (*)

#Akaike Information Criterion
## smaleer values = better fit model

AIC(model) # model with random effect of block
AIC(model_noRE) # model with no block

```


STEPS:
1. Calculate your variable
2. Add it to the table of diversity indives (whihc includes transect info)
3. Visualize it using box plot and/or histogram
4. Choose your random component
5. Compose your systematic component (your formula)
6. Decide if you have random effects to include
7. Choose which modeling function to use in R
- lm for normal fixed effects only
- glm for non-normal, fixed effects only
- lmer for normal with random effects
- glmer for non-normal with random effects
8. Plot(model) to see residual plot
9. Examine fixed effects estimates in summary(model)
10. Compare fit of model to other possible modeling using AIC()


#For assignment 
One for biodiversity, one for percent cover

1. Total Native hits/transect - poisson
2. Total Native hits/ Total Hits (per transect) --> proportion -  Beta/binomial
3. PerCoverENCCAL (Pre-abs/9) - binomial
4. 
5. 
6. 

