---
title: "Phase 2 -HL"
author: "Helen"
date: "2023-02-16"
output: html_document
---


```{r}

#Install Packages

library(tidyverse)
library(readxl)
library(janitor)
library(vegan)
library(lubridate)
library(dataspice)
library(lme4)
library(lmerTest)
```


## Formatting the data for calculation and GLM (percent vover)
```{r}
load(file = "DNTransect_Primary.Rdata")

DN_index <- DNTransect_Primary %>% 
  mutate(Index = paste(year(Date),Block,Plot,Transect, sep = "-"))

SpeciesPerCover <- DN_index %>%
  filter(Type == "Plant")%>%
  select(Index, SpeciesCode, Hits)%>%
  mutate(PresAbs = case_when(
    Hits >= 1 ~ 1,
    Hits < 1 ~ 0
    )) %>%
  group_by(Index, SpeciesCode)%>%
  summarise(PresAbsSum = sum(PresAbs))%>%
  mutate(PerCover = PresAbsSum/9) %>%
  pivot_wider(names_from = SpeciesCode, values_from = PerCover)%>%
  replace(is.na(.), 0)
  
```

```{r}
NativePerCover <- SpeciesPerCover %>%
  pivot_longer(3:ncol(SpeciesPerCover), names_to = "SpeciesCode", values_to = "PerCover")%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  select(-PresAbsSum)%>%
  filter(NativeStatus == "Native")%>%
  group_by(Index)%>%
  summarise(NativePerCover = sum(PerCover))

```


```{r}
DN_diversity_index_Native <- DN_diversity_index%>% left_join(NativePerCover)

ggplot(DN_diversity_index_Native)+
  aes(y = NativePerCover, x = Treatment, fill = Region)+
  geom_boxplot()
```

## creating histogram to visulaise data distribution
```{r}
DN_diversity_index_Native%>%
  filter(Region != "C")%>%
  ggplot()+
  aes(x = NativePerCover)+
  geom_histogram()+
  facet_grid(Treatment~Region)


# Three components of GLM
# 1. random component
# 2. systematic component
# 3. link function
library(lme4)
library(lmerTest)

test_1 <- DN_diversity_index_Native %>% 
  mutate (Year = year(Date))%>%
  filter(Region != "C", Year == 2022)

model <- lmer(NativePerCover~ Treatment * Region + (1| Block),
             data = test_1
             #family = "gaussian"
             )

model_noRE <- lm(NativePerCover~ Treatment * Region,  
                 data = test_1)
summary(model)

plot(model)

# To read the rsult -
##Estimate - 
###Treatment E and I (*) compared to ambient
###Region X compared to Region M (*)

#Akaike Information Criterion
## smaller values = better fit model

AIC(model) # model with random effect of block
AIC(model_noRE) # model with no block

```


STEPS:
1. Calculate your variable
2. Add it to the table of diversity indices (which includes transect info)
3. Visualize it using box plot and/or histogram
4. Choose your random component
5. Compose your systematic component (your formula)
6. Decide if you have random effects to include
7. Choose which modeling function to use in R
- lm for normal fixed effects only
- glm for non-normal, fixed effects only
- lmer for normal with random effects
- glmer for non-normal with random effects
8. Plot(model) to see residual plot
9. Examine fixed effects estimates in summary(model)
10. Compare fit of model to other possible modeling using AIC()


#For assignment 
One for biodiversity, one for percent cover

1. Total Native hits/transect - poisson
2. Total Native hits/ Total Hits (per transect) --> proportion -  Beta/binomial
3. PerCoverENCCAL (Pre-abs/9) - binomial
4. Total Non-native hits/transect - poisson
5. 
6. 

# Calculating the variable and the merge it with diversity index
```{r}
NonNative_hits <- DN_index %>%
  filter(Type == "Plant")%>%
  select(Index, SpeciesCode, Hits)%>%
  left_join(SpeciesKey, by = "SpeciesCode")%>%
  filter(NativeStatus == "Non-Native")%>%
  group_by(Index)%>%
  summarise(Non_Native_Hits = sum(Hits))%>%
  left_join(DN_diversity_index)

```

# Create histogram
```{r}
ggplot(NonNative_hits)+
  aes(y = Non_Native_Hits, x = Treatment, fill = Region)+
  geom_boxplot()

#NonNative <- NonNative_hits%>%
# filter(Region != "C")

ggplot(NonNative)+
  aes(x = Non_Native_Hits)+
  geom_histogram()+
  facet_grid(Treatment~Region)

ggplot(NonNative)+
  aes(x = Non_Native_Hits)+
  geom_histogram()
```

```{r}
NonNative_1 <- NonNative_hits %>% 
  mutate (Year = year(Date)) %>%
  filter(Region != "C" , Year == 2022)

model_nonnative <- glmer(Non_Native_Hits~ Treatment * Region + ( 1 + Treatment| Block), # systematic component
             data = NonNative_1, #specify data
             family = "gaussian" #poisson(link = "log") #random component for poisson distributed data
             )

model_poisson <- glmer (Non_Native_Hits~ Treatment * Region + ( 1+ Treatment | Block), 
             data = NonNative_1,
             family = poisson(link = "log")
                 )

plot(model_nonnative)
plot(model_poisson)

summary(model_nonnative)
summary(model_poisson)

AIC(model_nonnative) 
AIC(model_poisson) 
```

# for beta distribution class
```{r}
ENCCAL <- DN_index %>%
  filter(SpeciesCode == "ENCCAL")%>%
  select(Index, SpeciesCode, Hits)%>%
  mutate(PresAbs = case_when(
    Hits >= 1 ~ 1,
    Hits < 1 ~ 0))%>%
  group_by(Index)%>%
  summarise(ENCCALcover = sum(PresAbs)/9)%>%
  left_join(DN_diversity_index)


model_ENCCAL <- glm(ENCCALcover~ Treatment * Region, #+ ( 1| Block), # systematic component
             data = ENCCAL, #specify data
             family = "binomial" #random component for poisson distributed data
             )
plot(model_ENCCAL)
summary(model_ENCCAL)
AIC(model_ENCCAL)

```

