---
title: "Drought Net Data Cleaning"
author: "Amy Henry"
date: "2023-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Step 1: Install your packages
```{r}
install.packages("tidyverse")
install.packages("readxl")
install.packages("janitor")

library(tidyverse)
library(readxl)
library(janitor)
```

Step 2: Import your data
```{r}
## So far I have taught you to import data from CSV. This is not always practical - often data is in XLS or XLSX format. The readxl package helps with this. 

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx")
# multiple sheets mean we need to specify

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1")
# whoa loads of error messages - looking at the data, a lot is not formatted right, especially the date. 

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", skip = 1)
# ah, that didn't help, because we'd rather have kept the first line as the column names. 
# remove the skip. 

# Instead lets perhaps skip naming columns and just have everything come in as characters. 
Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)
```
There is more than one year's worth of data however. We need to import all three and check if they're in the same format. 
```{r}
Spring_2020_headers <- 
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

Spring_2021_headers <- 
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

Spring_2022_headers <- 
  read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

names(Spring_2020_headers)
names(Spring_2021_headers)
names(Spring_2022_headers)
```


Formatting one data frame - let's check out the checklist!

Which row has the best column names to keep?
Use colnames(data) <- data[rownumber, ] to name your columns after that row. 
```{r}

# the function row_to_names() is from the janitor package
Spring_2020_raw <- Spring_2020_raw %>% row_to_names(row_number(1))

Spring_2020_raw
```

Are those column names okay, or do they need to be changed to be tidy, contain identifying information? 
Use rename() to give your columns better names. 
```{r}
Spring_2020_raw <- Spring_2020_raw %>% rename(Transect = `Transect #`, Hits = `# Hits`)
```

Which rows don’t contain observations? 
Remove them from your dataset using [ , ] or filter().  
```{r}
Spring_2020_raw %>% filter(Date != "Date" & Date != "mm/dd/yy")
```
What should the columns be? 
- Date should be a in a date format
- team should be character
- block should be a factor and just say 1 instead of 1.0
- Plot should be a factor and just say 1 instead of 1.0
- Treatment should be a factor (and maybe will be more useful if it has the full word? exclusion etc.?)
- Region should be a factor
- Transect can be numeric or factor, doesn't really matter, but shouldn't have a decimal
- Distance should be numeric
- Code is character
- Hits should be numeric and not have a decimal
- Notes can stay character

 
string it together in one tidy pipe.  
```{r}
Spring_2020_raw <- 
  read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)  %>% 
  row_to_names(row_number(1)) %>% 
  rename(Transect = `Transect #`, Hits = `# Hits`) %>% 
  filter(Date != "mm/dd/yy")

```
 
 